---
title: "Sample Charts"
author: "Steve Mortimer"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: true
    code_folding: hide
    fig_width: 10
    fig_height: 10
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    theme: paper
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
---

# Introduction

The following chart examples require packages that you might not have currently
installed on your local machine. Please run this first chunk of code before
proceeding if you do not have these packages.

```{r install-packages, echo=T, eval=F}
install.packages('devtools')
install.packages('magrittr')
install.packages('xml2') ##### Work with XML files using a simple, consistent interface.
install.packages('purrr') ##### Functional programming tools for R
install.packages('plyr')
install.packages('dplyr')
install.packages('tidyr')
install.packages('stringr')
install.packages('knitr')
install.packages('readr')
install.packages('tsne') ##### t-distributed stochastic neighbor embedding
##### Wikipedia's description: It is a nonlinear dimensionality reduction technique
##### that is particularly well-suited for embedding high-dimensional data into a
##### space of two or three dimensions, which can then be visualized in a scatter # ##### plot.
install.packages('ggplot2')
install.packages('highcharter') ##### A wrapper for the 'Highcharts' library including shortcut functions to plot R objects.
##### 'Highcharts' <http://www.highcharts.com/> is a charting library offering
##### numerous chart types with a simple configuration syntax.
devtools::install_github('rstudio/htmltools') ##### Tools for HTML generation and output.
devtools::install_github('rstudio/leaflet')  ##### Leaflet is one of the most
##### popular open-source JavaScript libraries for interactive maps. 
##### R package for integrating and controling Leaflet maps in R.
devtools::install_github('rich-iannone/DiagrammeR')
```

``` {r global_parms, echo=FALSE, eval=TRUE, warning=FALSE, cache=FALSE}

options(stringsAsFactors=FALSE, scipen=999)

suppressWarnings(suppressPackageStartupMessages(library(knitr)))
knitr::opts_chunk$set(results='asis', echo=TRUE, eval=TRUE,
                      cache=FALSE, message = FALSE, warning = FALSE)

```

Load the packages necessary for doing some charting

```{r load-packages, echo=T}
suppressWarnings(suppressPackageStartupMessages(library(tsne)))
suppressWarnings(suppressPackageStartupMessages(library(plyr)))
suppressWarnings(suppressPackageStartupMessages(library(dplyr)))
suppressWarnings(suppressPackageStartupMessages(library(tidyr)))
suppressWarnings(suppressPackageStartupMessages(library(xml2)))
suppressWarnings(suppressPackageStartupMessages(library(readr)))
suppressWarnings(suppressPackageStartupMessages(library(purrr)))
suppressWarnings(suppressPackageStartupMessages(library(ggplot2)))
suppressWarnings(suppressPackageStartupMessages(library(highcharter)))
suppressWarnings(suppressPackageStartupMessages(library(htmltools)))
suppressWarnings(suppressPackageStartupMessages(library(stringr)))
suppressWarnings(suppressPackageStartupMessages(library(rvest)))
suppressWarnings(suppressPackageStartupMessages(library(leaflet)))
suppressWarnings(suppressPackageStartupMessages(library(DiagrammeR)))
suppressWarnings(suppressPackageStartupMessages(library(magrittr)))
```

# Pie Charts in `ggplot2`

Pie charts are generally frowned upon, but it's possible to make them in ggplot and the
way to create them is using polar coordinates, which is an interesting workaround
for a bar chart.

```{r pie-chart-data, echo=F}
iris_distribution <- as.data.frame(prop.table(table(iris$Species)))
names(iris_distribution) <- c('Species', 'Proportion')

```

```{r pie-chart-plot, echo=T}
ggplot(data=iris_distribution,
         aes(x=factor(1),
         y=Proportion,
         fill = factor(Species))) + 
 geom_bar(width = 1, colour="black", stat="identity") + 
 coord_polar(theta="y") +
 xlab('Proportion of Total') +
 ylab('') +
 labs(fill='Species')	+
 scale_x_discrete(breaks=NULL)+
 scale_y_continuous(breaks=seq(0, .75, 0.25),labels=c("100% / 0%","25%","50%","75%")) + 
 theme(panel.grid.major = element_blank())
```

# Interactive mapping with `leaflet`

Imagine you own Capital Bikeshare in DC. What is the shortest path to every single kiosk in the
city of Washington, DC? This problem can be solved by downloading the latitude and longitude
coordinates of each bike share location and then running an algorithm over those points
to determine the shortest path. Assuming you've loaded the TSP package and run the solution
here is an interesting visualization of the optimal path.

```{r load-leaflet-tsp-data, echo=F}
data <- structure(list(latitude = c(38.92469, 38.924646, 38.921263, 38.922168, 
38.918765, 38.916857, 38.905218, 38.905476, 38.903559, 38.901133, 
38.902078, 38.90149, 38.898837, 38.904986, 38.906431, 38.906435, 
38.896277, 38.896423, 38.893588, 38.885983, 38.884449, 38.883947, 
38.885531, 38.886919, 38.886669, 38.888853, 38.886216, 38.892595, 
38.892934, 38.897148, 38.897961, 38.896764, 38.891852, 38.888027, 
38.888301, 38.883832, 38.882348, 38.88073, 38.882752, 38.884181, 
38.882075, 38.870938, 38.874046, 38.87492, 38.87344, 38.866297, 
38.863403, 38.858684, 38.863313, 38.860892, 38.861129, 38.843038, 
38.838929, 38.835598, 38.829601, 38.827425, 38.833408, 38.837364, 
38.834783, 38.833856, 38.833648, 38.830615, 38.837071, 38.842278, 
38.849709, 38.845999, 38.855555, 38.857391, 38.869303, 38.869939, 
38.872104, 38.87351, 38.87587, 38.873638, 38.871136, 38.868376, 
38.88017, 38.881183, 38.883001, 38.885538, 38.885501, 38.89207, 
38.891676, 38.886206, 38.889225, 38.885703, 38.888698, 38.891268, 
38.893627, 38.902301, 38.90248, 38.903246, 38.902925, 38.903461, 
38.900086, 38.897321, 38.902678, 38.907002, 38.906193, 38.919268, 
38.921747, 38.930069, 38.929172, 38.931497, 38.934918, 38.934875, 
38.936715, 38.939416, 38.941603, 38.95008, 38.955897, 38.95659, 
38.955633, 38.956787, 38.960796, 38.962518, 38.96785, 38.968379, 
38.974502, 38.980954, 38.980572, 38.983378, 38.984523, 38.985747, 
38.98798, 38.980858, 38.976661, 38.971688, 38.967111, 38.966864, 
38.964767, 38.963521, 38.960559, 38.96098, 38.958061, 38.955217, 
38.955204, 38.954028, 38.952482, 38.95169, 38.948451, 38.948675, 
38.945128, 38.948053, 38.943689, 38.94197, 38.941506, 38.945442, 
38.945373, 38.94164, 38.939952, 38.932086, 38.932364, 38.930497, 
38.927824, 38.920329, 38.920371, 38.921344, 38.923367, 38.925589, 
38.928918, 38.930385, 38.930385, 38.93387, 38.934762, 38.93211, 
38.944474, 38.950393, 38.958927, 38.968968, 38.967822, 38.96562, 
38.965141, 38.968397, 38.962607, 38.953066, 38.956242, 38.95447, 
38.951808, 38.949498, 38.948993, 38.934471, 38.931537, 38.931946, 
38.925207, 38.922543, 38.923584, 38.91996, 38.917885, 38.927799, 
38.936486, 38.935688, 38.931635, 38.929518, 38.929345, 38.920851, 
38.918425, 38.918972, 38.917513, 38.91355, 38.909844, 38.90622, 
38.906434, 38.906955, 38.916704, 38.913915, 38.913814, 38.910956, 
38.909965, 38.913575, 38.914689, 38.916089, 38.916776, 38.915226, 
38.912859, 38.910569, 38.907719, 38.907379, 38.900493, 38.903168, 
38.903871, 38.905029, 38.908749, 38.91045, 38.910906, 38.912033, 
38.914653, 38.912813, 38.913221, 38.91529, 38.917247, 38.917933, 
38.920399, 38.921942, 38.925082, 38.928662, 38.929305, 38.929135, 
38.928912, 38.930892, 38.92891, 38.934249, 38.933307), longitude = c(-76.978702, 
-76.99059, -76.997534, -77.002985, -77.001814, -76.999135, -77.001496, 
-77.00005, -76.993221, -76.993063, -76.989093, -76.988621, -76.983047, 
-76.985906, -76.9806, -76.978719, -76.973281, -76.974288, -76.981154, 
-76.977936, -76.981989, -76.983447, -76.984412, -76.985745, -76.986616, 
-76.988634, -76.992408, -76.996372, -76.992413, -76.992395, -76.997333, 
-76.998661, -77.005538, -77.000781, -76.999742, -76.997361, -76.991819, 
-76.993533, -77.004064, -77.004803, -77.022412, -77.016675, -77.011539, 
-77.01159, -76.992401, -76.988457, -76.98242, -76.986088, -76.989413, 
-76.993419, -76.997192, -77.019541, -77.0219, -77.016727, -77.023423, 
-77.008389, -77.007389, -77.011409, -77.004059, -76.99795, -76.996361, 
-76.991721, -76.990945, -76.986051, -76.9842, -76.972549, -76.962725, 
-76.958568, -76.968034, -76.972432, -76.970186, -76.968042, -76.965204, 
-76.962678, -76.952488, -76.942321, -76.931564, -76.942278, -76.94034, 
-76.942666, -76.94327, -76.948542, -76.941061, -76.936297, -76.932455, 
-76.919518, -76.91947, -76.925797, -76.923993, -76.922726, -76.92515, 
-76.927256, -76.929915, -76.934427, -76.939458, -76.941893, -76.943012, 
-76.941122, -76.950338, -76.954753, -76.955459, -76.965304, -76.972423, 
-76.966477, -76.965717, -76.967353, -76.970934, -76.976961, -76.976511, 
-76.988523, -76.992291, -76.993437, -76.998161, -77.008228, -77.000789, 
-77.00209, -77.008818, -77.010484, -77.020733, -77.024163, -77.029239, 
-77.029954, -77.032555, -77.031797, -77.044239, -77.054722, -77.055576, 
-77.028209, -77.021462, -77.020829, -77.021035, -77.019691, -77.022671, 
-77.017786, -77.013406, -77.018456, -77.018935, -77.022373, -77.021892, 
-77.024296, -77.027837, -77.018831, -77.016727, -77.007096, -77.009815, 
-77.01687, -77.018949, -77.019768, -77.026788, -77.030424, -77.028917, 
-77.02481, -77.030634, -77.033913, -77.029118, -77.039613, -77.042655, 
-77.048025, -77.049161, -77.044202, -77.037902, -77.03959, -77.039661, 
-77.037394, -77.042165, -77.044458, -77.056286, -77.033153, -77.029959, 
-77.055342, -77.05587, -77.060719, -77.063211, -77.074348, -77.077438, 
-77.072046, -77.078104, -77.083262, -77.090849, -77.09362, -77.09976, 
-77.084305, -77.088115, -77.106502, -77.102848, -77.102561, -77.099968, 
-77.097714, -77.096921, -77.084738, -77.068053, -77.067285, -77.066023, 
-77.060069, -77.056653, -77.073338, -77.074441, -77.071495, -77.072033, 
-77.067488, -77.063495, -77.07043, -77.065048, -77.06345, -77.055444, 
-77.052149, -77.04852, -77.049479, -77.038268, -77.039239, -77.039419, 
-77.037732, -77.03051, -77.026685, -77.025277, -77.028744, -77.029879, 
-77.027693, -77.018321, -77.016452, -77.010345, -77.011678, -77.01634, 
-77.015458, -77.010031, -77.011169, -77.01283, -77.01553, -77.017982, 
-77.017438, -77.017682, -77.009993, -77.014547, -77.017931, -77.017308, 
-77.022949, -77.021286, -77.019964, -76.999274, -76.992335, -76.99104, 
-76.981012, -76.97988), popup = structure(c(159L, 27L, 8L, 214L, 
171L, 219L, 111L, 120L, 93L, 108L, 142L, 221L, 31L, 201L, 119L, 
186L, 217L, 195L, 44L, 60L, 155L, 70L, 32L, 17L, 188L, 86L, 106L, 
141L, 170L, 63L, 92L, 95L, 139L, 79L, 42L, 34L, 99L, 158L, 91L, 
168L, 20L, 207L, 72L, 97L, 232L, 191L, 222L, 193L, 96L, 238L, 
71L, 127L, 234L, 208L, 152L, 23L, 103L, 133L, 144L, 36L, 194L, 
206L, 198L, 57L, 5L, 180L, 123L, 230L, 160L, 239L, 242L, 102L, 
33L, 126L, 112L, 24L, 46L, 176L, 236L, 107L, 169L, 197L, 43L, 
74L, 184L, 233L, 211L, 26L, 146L, 69L, 224L, 151L, 178L, 35L, 
83L, 130L, 196L, 65L, 153L, 2L, 156L, 138L, 25L, 132L, 161L, 
213L, 66L, 210L, 68L, 173L, 147L, 1L, 59L, 89L, 243L, 179L, 166L, 
192L, 240L, 82L, 62L, 41L, 135L, 209L, 140L, 167L, 164L, 124L, 
199L, 14L, 143L, 175L, 216L, 149L, 125L, 134L, 73L, 88L, 77L, 
49L, 205L, 113L, 48L, 122L, 101L, 212L, 81L, 118L, 181L, 15L, 
47L, 165L, 18L, 80L, 235L, 241L, 94L, 84L, 10L, 229L, 7L, 226L, 
150L, 38L, 172L, 187L, 30L, 105L, 55L, 28L, 154L, 157L, 137L, 
64L, 116L, 58L, 237L, 76L, 90L, 189L, 39L, 19L, 131L, 223L, 29L, 
3L, 121L, 174L, 203L, 200L, 52L, 114L, 98L, 50L, 227L, 104L, 
13L, 225L, 45L, 53L, 100L, 40L, 190L, 204L, 202L, 162L, 78L, 
182L, 75L, 117L, 6L, 67L, 16L, 22L, 9L, 11L, 145L, 215L, 85L, 
37L, 136L, 54L, 12L, 87L, 4L, 109L, 185L, 21L, 51L, 177L, 220L, 
128L, 115L, 56L, 129L, 61L, 183L, 231L, 218L, 228L, 148L, 163L, 
110L), .Label = c("Kiosk Location #1", "Kiosk Location #10", 
"Kiosk Location #100", "Kiosk Location #101", "Kiosk Location #102", 
"Kiosk Location #103", "Kiosk Location #104", "Kiosk Location #105", 
"Kiosk Location #106", "Kiosk Location #107", "Kiosk Location #108", 
"Kiosk Location #109", "Kiosk Location #11", "Kiosk Location #110", 
"Kiosk Location #111", "Kiosk Location #112", "Kiosk Location #113", 
"Kiosk Location #114", "Kiosk Location #115", "Kiosk Location #116", 
"Kiosk Location #117", "Kiosk Location #118", "Kiosk Location #119", 
"Kiosk Location #12", "Kiosk Location #120", "Kiosk Location #121", 
"Kiosk Location #122", "Kiosk Location #123", "Kiosk Location #124", 
"Kiosk Location #125", "Kiosk Location #126", "Kiosk Location #127", 
"Kiosk Location #128", "Kiosk Location #129", "Kiosk Location #13", 
"Kiosk Location #130", "Kiosk Location #131", "Kiosk Location #132", 
"Kiosk Location #133", "Kiosk Location #134", "Kiosk Location #135", 
"Kiosk Location #136", "Kiosk Location #137", "Kiosk Location #138", 
"Kiosk Location #139", "Kiosk Location #14", "Kiosk Location #141", 
"Kiosk Location #142", "Kiosk Location #143", "Kiosk Location #144", 
"Kiosk Location #145", "Kiosk Location #146", "Kiosk Location #147", 
"Kiosk Location #148", "Kiosk Location #149", "Kiosk Location #15", 
"Kiosk Location #150", "Kiosk Location #151", "Kiosk Location #152", 
"Kiosk Location #153", "Kiosk Location #154", "Kiosk Location #155", 
"Kiosk Location #156", "Kiosk Location #157", "Kiosk Location #158", 
"Kiosk Location #159", "Kiosk Location #16", "Kiosk Location #160", 
"Kiosk Location #161", "Kiosk Location #162", "Kiosk Location #163", 
"Kiosk Location #164", "Kiosk Location #165", "Kiosk Location #166", 
"Kiosk Location #167", "Kiosk Location #168", "Kiosk Location #169", 
"Kiosk Location #17", "Kiosk Location #170", "Kiosk Location #171", 
"Kiosk Location #172", "Kiosk Location #173", "Kiosk Location #174", 
"Kiosk Location #175", "Kiosk Location #176", "Kiosk Location #177", 
"Kiosk Location #178", "Kiosk Location #179", "Kiosk Location #18", 
"Kiosk Location #180", "Kiosk Location #181", "Kiosk Location #182", 
"Kiosk Location #183", "Kiosk Location #184", "Kiosk Location #185", 
"Kiosk Location #186", "Kiosk Location #187", "Kiosk Location #188", 
"Kiosk Location #189", "Kiosk Location #19", "Kiosk Location #190", 
"Kiosk Location #191", "Kiosk Location #192", "Kiosk Location #193", 
"Kiosk Location #194", "Kiosk Location #195", "Kiosk Location #196", 
"Kiosk Location #197", "Kiosk Location #198", "Kiosk Location #199", 
"Kiosk Location #2", "Kiosk Location #20", "Kiosk Location #200", 
"Kiosk Location #201", "Kiosk Location #202", "Kiosk Location #203", 
"Kiosk Location #204", "Kiosk Location #205", "Kiosk Location #206", 
"Kiosk Location #207", "Kiosk Location #208", "Kiosk Location #209", 
"Kiosk Location #21", "Kiosk Location #210", "Kiosk Location #211", 
"Kiosk Location #212", "Kiosk Location #213", "Kiosk Location #214", 
"Kiosk Location #215", "Kiosk Location #216", "Kiosk Location #217", 
"Kiosk Location #218", "Kiosk Location #219", "Kiosk Location #22", 
"Kiosk Location #220", "Kiosk Location #221", "Kiosk Location #222", 
"Kiosk Location #223", "Kiosk Location #224", "Kiosk Location #225", 
"Kiosk Location #226", "Kiosk Location #227", "Kiosk Location #228", 
"Kiosk Location #229", "Kiosk Location #23", "Kiosk Location #230", 
"Kiosk Location #231", "Kiosk Location #232", "Kiosk Location #233", 
"Kiosk Location #234", "Kiosk Location #235", "Kiosk Location #236", 
"Kiosk Location #237", "Kiosk Location #238", "Kiosk Location #239", 
"Kiosk Location #24", "Kiosk Location #240", "Kiosk Location #241", 
"Kiosk Location #242", "Kiosk Location #243", "Kiosk Location #244", 
"Kiosk Location #25", "Kiosk Location #26", "Kiosk Location #27", 
"Kiosk Location #28", "Kiosk Location #29", "Kiosk Location #3", 
"Kiosk Location #30", "Kiosk Location #31", "Kiosk Location #32", 
"Kiosk Location #33", "Kiosk Location #34", "Kiosk Location #35", 
"Kiosk Location #36", "Kiosk Location #37", "Kiosk Location #38", 
"Kiosk Location #39", "Kiosk Location #4", "Kiosk Location #40", 
"Kiosk Location #41", "Kiosk Location #42", "Kiosk Location #43", 
"Kiosk Location #44", "Kiosk Location #45", "Kiosk Location #46", 
"Kiosk Location #47", "Kiosk Location #48", "Kiosk Location #49", 
"Kiosk Location #5", "Kiosk Location #50", "Kiosk Location #51", 
"Kiosk Location #52", "Kiosk Location #53", "Kiosk Location #54", 
"Kiosk Location #55", "Kiosk Location #56", "Kiosk Location #57", 
"Kiosk Location #58", "Kiosk Location #59", "Kiosk Location #6", 
"Kiosk Location #60", "Kiosk Location #61", "Kiosk Location #62", 
"Kiosk Location #63", "Kiosk Location #64", "Kiosk Location #65", 
"Kiosk Location #66", "Kiosk Location #67", "Kiosk Location #68", 
"Kiosk Location #69", "Kiosk Location #7", "Kiosk Location #70", 
"Kiosk Location #71", "Kiosk Location #72", "Kiosk Location #73", 
"Kiosk Location #74", "Kiosk Location #75", "Kiosk Location #76", 
"Kiosk Location #77", "Kiosk Location #78", "Kiosk Location #79", 
"Kiosk Location #8", "Kiosk Location #80", "Kiosk Location #81", 
"Kiosk Location #82", "Kiosk Location #83", "Kiosk Location #84", 
"Kiosk Location #85", "Kiosk Location #86", "Kiosk Location #87", 
"Kiosk Location #88", "Kiosk Location #89", "Kiosk Location #9", 
"Kiosk Location #90", "Kiosk Location #91", "Kiosk Location #92", 
"Kiosk Location #93", "Kiosk Location #94", "Kiosk Location #95", 
"Kiosk Location #96", "Kiosk Location #97", "Kiosk Location #98", 
"Kiosk Location #99"), class = "factor")), .Names = c("latitude", 
"longitude", "popup"), class = "data.frame", row.names = c(NA, 
-243L))

```

```{r plot-leaflet-tsp, echo=T}
map <- leaflet() %>%
  addProviderTiles("Stamen.TonerLite",
                   options = providerTileOptions(noWrap = TRUE)) %>%
  addCircleMarkers(data=data, lat = ~latitude, lng = ~longitude, 
                   radius = 3, popup = ~as.character(paste0(latitude,",",longitude))) %>%
  addPolylines(data=data, lng = ~longitude, lat = ~latitude, 
               color = "#A93E36", opacity = .7)
map
```

# Using `diagrammeR` to plot graph data

Imagine you have a series of interconnected tasks and processes. If you'd like to
visualize them using R, then you can use the diagrammeR package

```{r create-diagram-visualization, echo=T}
graph <-
  create_graph() %>%
  set_graph_name("DAG") %>%
  set_global_graph_attrs("graph", "overlap", "true") %>%
  set_global_graph_attrs("graph", "fixedsize", "true") %>%
  set_global_graph_attrs("node", "color", "blue") %>%
  set_global_graph_attrs("node", "fontname", "Helvetica") %>%
  add_n_nodes(11) %>%
  select_nodes_by_id(c(1:4, 8:11)) %>% 
  set_node_attrs_ws("shape", "box") %>%
  clear_selection %>%
  select_nodes_by_id(5:7) %>% 
  set_node_attrs_ws("shape", "circle") %>%
  clear_selection %>%
  add_edges_w_string(
    "1->5 2->6 3->9 4->7 5->8 5->10 7->11", "green") %>%
  add_edges_w_string(
    "1->8 3->6 3->11 3->7 5->9 6->10", "red") %>%
  select_edges("rel", "green") %>%
  set_edge_attrs_ws("color", "green") %>%
  invert_selection %>%
  set_edge_attrs_ws("color", "red")

render_graph(graph)
```

# The Pokeman plot

If you love pokemon as much as the person who wrote this ["Visualize 'Em All"](http://jkunst.com/r/pokemon-visualize-em-all/) 
blog post then you might find their visualization just ridiculous.

```{r loading-pokemon-data, echo=F, eval=T, cache=TRUE}
path <- function(x) paste0("https://raw.githubusercontent.com/phalt/pokeapi/master/data/v2/csv/", x)

dfpkmn <- read_csv(path("pokemon.csv")) %>%
  select(-order, -is_default) %>%
  dplyr::rename(pokemon = identifier) 

# dfpkmn <- read_csv("https://raw.githubusercontent.com/phalt/pokeapi/master/data/v2/csv/pokemon.csv") %>% 
#   select(-order, -is_default) %>% 
#   mutate(pokemon = identifier) %>%
#   select(-identifier)


dfstat <- read_csv(path("stats.csv")) %>% 
  dplyr::rename(stat_id = id) %>%
  right_join(read_csv(path("pokemon_stats.csv")),
             by = "stat_id") %>% 
  dplyr::mutate(identifier = str_replace(identifier, "-", "_")) %>% 
  select(pokemon_id, identifier, base_stat) %>% 
  spread(identifier, base_stat) %>% 
  dplyr::rename(id = pokemon_id)

dftype <- read_csv(path("types.csv")) %>% 
  dplyr::rename(type_id = id) %>%
  right_join(read_csv(path("pokemon_types.csv")), by = "type_id") %>% 
  select(pokemon_id, identifier, slot) %>% 
  mutate(slot = paste0("type_", slot)) %>% 
  spread(slot, identifier) %>% 
  dplyr::rename(id = pokemon_id) 

dfegg <- read_csv(path("egg_groups.csv")) %>% 
  dplyr::rename(egg_group_id = id) %>% 
  right_join(read_csv(path("pokemon_egg_groups.csv")), by = "egg_group_id") %>% 
  group_by(species_id) %>% 
  dplyr::mutate(ranking = row_number(),
         ranking = paste0("egg_group_", ranking)) %>% 
  select(species_id, ranking, identifier) %>% 
  spread(ranking, identifier) 

dfimg <- "https://github.com/phalt/pokeapi/tree/master/data/Pokemon_XY_Sprites" %>% 
  read_html() %>% 
  html_nodes("tr.js-navigation-item > .content > .css-truncate a") %>% 
  map_df(function(x){
    url <- x %>% html_attr("href")
    data_frame(
      id = str_extract(basename(url), "\\d+"),
      url_image = basename(url)
    )
  }) %>%
  mutate(id = as.numeric(id))

url_bulbapedia_list <- "http://bulbapedia.bulbagarden.net/wiki/List_of_Pok%C3%A9mon_by_base_stats_(Generation_VI-present)" 

suppressWarnings(id <- url_bulbapedia_list %>% 
  read_html(encoding = "UTF-8") %>% 
  html_node("table.sortable") %>% 
  html_table() %>% 
  .[[1]] %>% 
  as.numeric())

url_icon <-  url_bulbapedia_list %>% 
  read_html() %>%
  html_nodes("table.sortable img") %>% 
  html_attr("src")

dficon <- data_frame(id, url_icon) %>% 
  dplyr::filter(!is.na(id)) %>% 
  dplyr::distinct(id, .keep_all = TRUE)

dfcolor <- map_df(na.omit(unique(c(dftype$type_1, dftype$type_2))), function(t){
  # t <- "bug"
  col <- "http://pokemon-uranium.wikia.com/wiki/Template:%s_color" %>% 
    sprintf(t) %>%
    read_html() %>% 
    html_nodes("span > b") %>% 
    html_text()
  data_frame(type = t, color = paste0("#", col))
})

dfcolorf <- expand.grid(color_1 = dfcolor$color, color_2 = dfcolor$color,
                        stringsAsFactors = FALSE) %>% 
  tbl_df() %>% 
  group_by(color_1, color_2) %>% 
  do({
      n = 100;p = 0.25
      data_frame(color_f = colorRampPalette(c(.$color_1, .$color_2))(n)[round(n*p)])
    })

# THE join
df <- dfpkmn %>% 
  left_join(dftype, by = "id") %>% 
  left_join(dfstat, by = "id") %>% 
  left_join(dfcolor %>% dplyr::rename(type_1 = type, color_1 = color), by = "type_1") %>% 
  left_join(dfcolor %>% dplyr::rename(type_2 = type, color_2 = color), by = "type_2") %>% 
  left_join(dfcolorf, by =  c("color_1", "color_2")) %>% 
  left_join(dfegg, by = "species_id") %>% 
  left_join(dfimg, by = "id") %>% 
  left_join(dficon, by = "id")

rm(dftype, dfstat, dfcolor, dfcolorf, dfegg, dfimg, dficon)
rm(id, url_bulbapedia_list, url_icon)

set.seed(13242)

tsne_poke <- df %>% 
  select(type_1, type_2, weight, height, base_experience,
         attack, defense, special_attack, special_defense, speed, base_experience,
         hp, egg_group_1, egg_group_2) %>%
  map(function(x){
    ifelse(is.na(x), "NA", x)
  }) %>% 
  as.data.frame() %>% 
  tbl_df() %>% 
  model.matrix(~., data = .) %>% 
  as.data.frame() %>% 
  tbl_df() %>% 
  .[-1] %>% 
  tsne(perplexity = 60)

df <- df %>% 
  dplyr::mutate(x = tsne_poke[, 1],
         y = tsne_poke[, 2])

dfcenters <- df %>% 
  group_by(type_1, color_1) %>% 
  summarise(cx = mean(x),
            cy = mean(y),
            sdcx = sd(x),
            sdcy = sd(y))
```

```{r plotting-pokemon, echo=T, eval=T}

ds <- df %>% 
  select(pokemon, type_1, type_2, weight, height,
         attack, defense, special_attack, special_defense,
         url_image, url_icon, color = color_1, x, y) %>% 
  list_parse() %>% 
  map(function(x){
    x$marker$symbol <- sprintf("url(%s)", x$url_icon)
    x$marker$radius <- 2
    x$url_icon  <- NULL
    x
  })

ds2 <- df %>% 
  select(color = color_1, x, y) %>%
  mutate(color = hex_to_rgba(color, 0.05)) %>% 
  list_parse()

urlimage <- "https://raw.githubusercontent.com/phalt/pokeapi/master/data/Pokemon_XY_Sprites/"

tooltip <- c("pokemon", "type_1", "type_2",
             "weight", "height",
             "attack",  "defense",
             "special_attack", "special_defense") %>%
  map(function(x){
    tags$tr(
      tags$th(str_replace_all(str_to_title(x), "_", " ")),
      tags$td(paste0("{point.", x, "}"))
    )
  }) %>% 
  do.call(tagList, .) %>% 
  tagList(
    tags$img(src = paste0(urlimage, "{point.url_image}"),
             width = "125px", height = "125px")
  ) %>% 
  as.character()

hctsne <- highchart() %>% 
  hc_chart(zoomType = "xy") %>% 
  hc_xAxis(minRange = diff(range(df$x))/5) %>%
  hc_yAxis(minRange = diff(range(df$y))/5) %>%
  hc_add_series(data = ds,
                type = "scatter",
                name = "pokemons",
                states = list(hover = list(halo = list(
                  size  = 50,
                  attributes = list(
                    opacity = 1)
                )))) %>%
  hc_add_series(data = ds2, type = "scatter",
                marker = list(radius = 75, symbol = "circle"),
                zIndex = -3,  enableMouseTracking = FALSE,
                linkedTo = ":previous") %>%
  hc_plotOptions(series = list()) %>%  
  hc_tooltip(
    useHTML = TRUE,
    borderRadius = 0,
    borderWidth = 5,
    headerFormat = "<table>",
    pointFormat = tooltip,
    footerFormat = "</table>"
  ) %>% 
  hc_add_theme(
    hc_theme_null(
      chart = list(
        backgroundColor = "transparent",
        style = list(
          fontFamily = "Roboto"
          )
        )
      )
    )

hctsne
```
